### 2021-11-19 README.Rmd
Update README with fancy instructions

# Abstract

@@

This repository contains the data and code for our paper:

> `r Authors`. (`r Year`). *`r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`.

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## How to cite

Please cite this compendium as:

> `r Authors`. (`r format(Sys.Date(), "%Y")`). *Compendium of R code and data for `r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`.

### How to download or install

@@

In order to reproduce this project, you'll need to install ASDF:

### `r emoji::emoji_glue(":warning::bird:")`: "Cautioning canary" project name

@@

### Licenses

@@

### Contributions

>>>>

[![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FMTR6X-blue)](https://dx.doi.org/10.17605/OSF.IO/MTR6X) [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4058987.svg)](https://doi.org/10.5281/zenodo.4058987)

> `r Authors`. `r Year`. ["`r Title`,"](https://doi.org/10.1177/0899764020971045) *Journal of Human Rights* (forthcoming).

**All this project's materials are free and open:**

- [Download the data](#data)
- [See the analysis notebook website](https://stats.andrewheiss.com/cautioning-canary/)

![Open data](img/data_large_color.png) &emsp; ![Open](img/materials_large_color.png)

## Abstract

@@

This repository contains the data and code for our paper. Our pre-print is online here:

> `r Authors`. `r Year`. "`r Title`"". Accessed `r format(Sys.time(), '%B %e, %Y')`. Online at <https://dx.doi.org/10.17605/OSF.IO/MTR6X>

## How to download and replicate

@@

We use the [**renv** package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [**targets** package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis. ([See this for a short helpful walkthrough of **targets**.](https://books.ropensci.org/targets/walkthrough.html)).

To reproduce the findings and re-run the analysis, do the following:

1. Download and install these fonts (if you’re using Windows, make sure you right click on the font files and choose “Install for all users” when installing these fonts):
    - [Cochineal](https://fontesk.com/cochineal-typeface/)
    - [Inter](https://fonts.google.com/specimen/Inter)
    - [Linux Libertine O](https://www.cufonfonts.com/font/linux-libertine-o) (also [here](https://sourceforge.net/projects/linuxlibertine/))
    - [Libertinus Math](https://github.com/alerque/libertinus)
    - [InconsolataGo](https://github.com/ryanoasis/nerd-fonts/tree/master/patched-fonts/InconsolataGo)
2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
    - If you're using macOS, [install XQuartz too](https://www.xquartz.org/), so that you have access to the Cairo graphics library
    - If you’re using Windows, [install RTools too](https://cran.r-project.org/bin/windows/Rtools/) and add it to your PATH so that you can install packages from source if needed
3. Open `cautioning-canary.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
4. Make sure you have a working installation of LaTeX:
    - *Easy-and-recommended way*: Install the [**tinytex** package](https://yihui.org/tinytex/) by running `install.packages("tinytex")` in the R console, then running `tinytex::install_tinytex()`
    - *Easy-but-requires-huge-4+-GB-download way*: Download TeX Live ([macOS](http://www.tug.org/mactex/); [Windows](https://miktex.org/))
5. If it's not installed already, R *should* try to install the **renv** package when you open the RStudio Project for the first time. If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.
6. Run `renv::restore()` in the R console to install all the required packages for this project.
7. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.

Running `targets::tar_make()` will create several helpful outputs:

1. All project data in `data/`
2. An analysis notebook website in `analysis/_site/index.html`
3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/`

## Data

<div style="color:#FF4136">
**NB**: If you're reproducing this project, it is best if you rely on the **targets** package and all the functions in `R/funs_data-cleaning.R` to handle the cleaning, processing, tidying, and merging. All the analysis in the project depends on having a **targets**-created object named `panel`.
</div>

For reference, though, we export CSV and RDS files of the final dataset:

- [**`data/derived_data/panel.csv`**](data/derived_data/panel.csv) and [**`data/derived_data/panel.rds`**](data/derived_data/panel.rds): CSV and RDS versions of our final combined dataset
- [**`data/derived_data/panel_lagged.csv`**](data/derived_data/panel_lagged.csv) and [**`data/derived_data/panel_lagged.rds`**](data/derived_data/panel_lagged.rds): CSV and RDS versions of our final combined dataset with lagged and leaded versions of variables

This data is derived from many different original data sources. See the ["Process and merge data" page of our analysis notebook](https://stats.andrewheiss.com/cautioning-canary/01_data-overview.html) for details about both the original data and the merging process.

- **Chaudhry NGO restrictions**: We use counts of anti-NGO legal barriers from [the replication data](https://doi.org/10.7910/DVN/JHOGNX) for Suparna Chaudhry's "The Assault on Civil Society: Explaining State Repression of NGOs" (*International Organization*, 2022).
    - `data/raw_data/Chaudhry restrictions/SC_Expanded.dta`
- **Political Terror Scores**: We use data from the [Political Terror Scale (PTS) project](http://www.politicalterrorscale.org/) to measure state repression. This project uses reports from the US State Department, Amnesty International, and Human Rights Watch and codes political repression on a scale of 1-5.
    - `data/raw_data/Political Terror Scale/PTS-2019.RData`, v2019
- **Latent Human Rights Protection Scores**: We use Chris Fariss's [Latent Human Rights Protection Scores](https://doi.org/10.7910/DVN/RQ85GK), which are estimates from fancy Bayesian models that capture a country's respect for physical integrity rights.
    - `data/raw_data/Latent Human Rights Protection Scores/HumanRightsProtectionScores_v4.01.csv`, v4.01
- **Varieties of Democracy data**: We use a bunch of variables from the [Varieties of Democracy (V-Dem) project](https://www.v-dem.net/en/).
    - `data/raw_data/Country_Year_V-Dem_Full+others_R_v10/V-Dem-CY-Full+Others-v10.rds`, v10
- **UN data**: We use [the **WDI** package](https://vincentarelbundock.github.io/WDI/) to collect data from the World Bank. However, we don't use WDI data for GDP and % of GDP from trade because the WDI data is incomplete (especially pre-1990, but that's not an issue in this project) To get around that, we create our own GDP and trade measures using data directly from the UN (at [UNData](https://data.un.org/)). They don't have a neat API like the World Bank, so you have to go to their website and export the data manually. We collect three variables: 
    - [GDP at constant 2015 prices](http://data.un.org/Data.aspx?q=gdp&d=SNAAMA&f=grID%3a102%3bcurrID%3aUSD%3bpcFlag%3a0) (`data/raw_data/UN data/UNdata_Export_20210118_034054729.csv`)
    - [GDP at current prices](http://data.un.org/Data.aspx?q=gdp&d=SNAAMA&f=grID%3a101%3bcurrID%3aUSD%3bpcFlag%3a0) (`data/raw_data/UN data/UNdata_Export_20210118_034311252.csv`)
    - [Population](https://population.un.org/wpp/Download/Standard/Population/) (`data/raw_data/UN data/WPP2019_POP_F01_1_TOTAL_POPULATION_BOTH_SEXES.xlsx`)
- **UCDP/PRIO Armed Conflict Data**: We use [UCDP/PRIO Armed Conflict data](https://ucdp.uu.se/downloads/index.html#armedconflict) to create an indicator marking if a country-year was involved in armed conflict that resulted in at least 25 battle-related deaths.
    - `data/raw_data/UCDP PRIO/ucdp-prio-acd-191.csv`, v19.1
- **Natural Earth shapefiles**: We use the ["Admin 0 - Countries" 1:110m cultural shapefiles](https://www.naturalearthdata.com/downloads/110m-cultural-vectors/) for maps.
    - `data/raw_data/ne_110m_admin_0_countries/`
- **2020 Civicus Monitor ratings**: CIVICUS rates countries [using a 5-item scale of civic space openness](https://monitor.civicus.org/widgets/world/), but getting their data in a machine-readable format is a little tricky. We downloaded [the standalone embeddable widget](https://monitor.civicus.org/widgets/world/) as an HTML file with `wget https://monitor.civicus.org/widgets/world/` and saved it as `index_2021-03-19.html`. We then extracted the `COUNTRIES_DATA` variable embedded in a `<script>` tag (`xpath = /html/body/script[5]`), which is JSON-ish, but not quite. The **jsonlite** R package couldn't parse it for whatever reason, but some random online JSON formatter and validator could, so we ran it through that and saved the resulting clean file.
    - `data/raw_data/Civicus/index_2021-03-19.html`
    - `data/raw_data/Civicus/civicus_2021-03-19.json`

**BUT AGAIN** we *highly* recommend just relying on the **targets** pipeline to construct this data rather than doing it all yourself! ([See this for a short helpful walkthrough.](https://books.ropensci.org/targets/walkthrough.html))

## `r emoji::emoji_glue(":warning::bird:")`: "Cautioning canary" project name

@@

## Licenses

@@

## Contributions
